Global Warming Distributed Data Analysis

### 2016-03-28 report/article.tex
conclusions

\section{Research Question}
\label{sec:question}
Our aim is to update the recent work done in the climate analysis with the new data available. We will first analyze the structure and characteristics of the data provided by the NOAA organization.
The data have a very important size and furthermore the data are continuously growing in size also in the last years. So one important question is \textbf{how to efficiently perform global analysis on huge wheater data?} We will perform the analysis utilizing distributed computing and study the advantages and disadvantages of this technology.\\
Then we are going to perform different analysis on the temperature data and compare the results with the ones found in literature. So it's important to find \textbf{which are the metrics that show better a change in the temperature?} In the end, we will make an evaluation on the threat of global warming based on the results we obtained, examining the utility of the selected metrics. \\

@@

@@

>>>>

\section{Research Question}
\label{sec:question}
Our aim is to update the recent work done in the climate analysis with the new data available. We will first analyze the structure and characteristics of the data provided by the NOAA organization.
The data have a very important size and furthermore the data are continuously growing in size also in the last years. So one important question is \textbf{how to efficiently perform global analysis on huge weather data?} We will perform the analysis utilizing distributed computing and study the advantages and disadvantages of this technology.\\
Then we are going to perform different analysis on the temperature data and compare the results with the ones found in literature. So it's important to find \textbf{which are the metrics that show better a change in the temperature?} In the end, we will make an evaluation on the threat of global warming based on the results we obtained, examining the utility of the selected metrics. \\

@@

\subsubsection{Standard Deviation analysis}
\begin{figure}[tbh]
\includegraphics[width=1\linewidth]{stdev}
\caption{Standard deviation for Average, Minimum and Maximum values per region.}
\label{fig:stdev}
\end{figure}
In figure \ref{fig:stdev} we plotted the value of the standard deviation per statistics and per region. It's clear that the values of maximum and minimum have larger fluctuation than the average. This may be caused by one or a few erroneous data report caused by environmental accidents. Interesting to note that the minimum value have a way larger standard deviation that the maximum. By correlating this with the stations distribution we came to the conclusion that this is caused by the installation of new stations in more remote and difficult to access regions (e.g. Arctic ocean and mountain regions). \\
Moreover, the minimum standard deviation for the Australian region was suspect, because the lowest temperature registered for the region is -23.4\degree C\cite{ausTemp}, so it's difficult to expect a standard deviation similar to that of the European region or even higher that that of the American one. After analyzing the data we found out erroneous data coming from the south east region of Australia reporting temperature of around -70\degree C.

@@

\subsection{Global Warming remarks}
From the worldwide analysis and using the interactive map, it appears that a clear temperature increase is in place around the world. Moreover, we notice this phenomenon occurring even considering specific large regions of earth where two out of three show a clear temperature increase trend for the last 30 years, which is in accordance with the data reported by independent researches\cite{hansen2006global}. \\
What we miss in the picture is the data coming from most of the oceanic areas and in general ocean temperatures. The only sources we had that covered large part of the oceans were coming from mobile stations installed on ships. The problem with this data was dual: many times we had the results (average, minimum and maximum) for a region constituted by only one station report. In the end we excluded these from the analysis because they skewed the results in an unpredictable way.

\subsection{Future Improvements}
Due to the limited amount of time available, there are many improvements to the analysis and other research path to discover for a better understanding of the dataset and the global warming phenomenon. \\
To improve the data quality we could divide the temperature data by time of day. We assumed that all the station have a similar behavior (they collect temperature samples every fixed amount of time). Without this assumption the average temperature result could lean towards lower or higher temperatures depending on which the hours for which we have more data.\\
Another interesting analysis that can be performed is dividing the regions by latitude bands. We expect that regions in the same latitude have lower differences in temperature so we could expect a more stable average temperature and less fluctuations in minimum/maximum values. The same reasoning can be done for seasons. In particular it would be particularly interesting to see the temperature changes by season for a fixed region and compare the season analysis (e.g. see if all seasons show the same curve or not).\\ 
An intriguing data source we had available but didn't use yet is the Arctic ice surface coverage. Since this is directly correlated with global temperature and is also one of the most monitored metric in recent years (due to its effects on the ecosystem), we think it can help in strengthening our conclusions. We didn't investigate further this path because our understanding of the data sources was limited and required more time than we expected to interpret correctly the data.
